---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(purler)

suppressPackageStartupMessages({
library(tidyr)
library(bench)
library(dplyr)
library(ggplot2)
})
```

# purler


<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
![](https://img.shields.io/badge/dependencies-zero-blue.svg)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![R build status](https://github.com/coolbutuseless/purler/workflows/R-CMD-check/badge.svg)](https://github.com/coolbutuseless/purler/actions)
<!-- badges: end -->


`purler` contains tools for run-length encoding vector data.

#### Key features:

* `NA` values are considered identical (unlike `base::rle()`)
* Faster calculation with results returned as a `data.frame`


## What's in the box

* `rlenc()` is C code for run-length encoding of raw, logical, integer, numeric and character vectors. 
    * Groups `NA` values into a run (unlike `base::rle()`)
    * Returns a **data.frame** rather than a list
    * Returned object is compatible with `base::inverse.rle()`
    * Can be **10x** faster than `base::rle()`
* `rle2()` is more compatible with `base::rle()` for values that `rlenc()` does not handle.
    * Written in vanilla R, so speed is about the same as `base::rle()`
    * Groups `NA` values into a run (unlike `base::rle()`)
    * Same type of return object as `base::rle()`
    * Will work with anything that `base::rle()` works with
* `rleid()` returns an integer vector numbering the runs of identical values within a
  vector of numeric or character data.  This is very similar to `data.table::rleid()`, 
  execpt the `data.table()` version is much more configurable and flexible.  This
  version is probably only useful if you wanted to avoid pulling in `data.table` as a
  dependency.

## Installation

You can install from [GitHub](https://github.com/coolbutuseless/purler) with:

``` r
# install.package('remotes')
remotes::install_github('coolbutuseless/purler')
```


## ToDo

* Long vector support




# `rlenc()` - run-length encoding output as a data.frame

```{r}
input <- c(1, 1, 1, 2, 2, 8, 8, 8, 8, 8, NA, NA, NA, NA)

(result <- purler::rlenc(input))

inverse.rle(result)
```


# `rlenc()` benchmark

```{r}
library(tidyr)
library(bench)
library(dplyr)
library(ggplot2)

N <- 1000
M <- 10

zz <- sample(seq_len(M), N, replace = TRUE)

res <- bench::mark(
  rle(zz),
  rlenc(zz),
  check = FALSE
)

plot(res)
```





# Run-length encoding with NAs

In `base::rle()`, runs of NA values are *not* treated as a group.

All functions in `purler` do treat NAs as identical for the purpose of creating groups

```{r}
input <- c(1, 1, 2, NA, NA, NA, NA, 4, 4, 4)

base::rle(input)

purler::rle2(input)

purler::rlenc(input)

purler::rlenc_id(input)
```





# Run-length encoded group ids

`rlenc_id()` numbers the runs of identical values in a numeric or character vector.

For a more complete approach to this problem, see `data.table::rleid()`

```{r}
input <- c(11, 11, 12, 12, 12, NA, NA, NA, NA)

rlenc_id(input)
```



## Related Software

* `base::rle()`
* `data.table::rleid()`


## Acknowledgements

* R Core for developing and maintaining the language.
* CRAN maintainers, for patiently shepherding packages onto CRAN and maintaining
  the repository
